A Compound Decision Approach to Covariance Matrix Estimation

### 2021-12-07 msg.tex
small changes for manuscript and response of reviews

\begin{prop}
  \label{prop:Rhat}
  Define $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ and
  \begin{equation}
    \label{eq:Rhat}
    \hat{R}_L(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ].
  \end{equation}
  If $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$, then
  \[
    \lim_{n \rightarrow 0} \{ \bb{E} \hat{R}_L(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta}) \} = 0
  \]
  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}.
\end{prop}
The condition on $\bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ requires that the variances of the entries of $\bs{S}$ do not grow too quickly as $n$ grows. It is also required by \citet{ledoit2004well} and is implied by their Lemma 3.1.

Proposition \ref{prop:Rhat} shows that $\hat{R}_L(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}. It is then reasonable to estimate the best estimator in this class by minimizing $\hat{R}_L(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. A slight modification of this procedure turns out to recover the \citet{ledoit2004well} estimator
\begin{equation}
\label{eq:lw}
\hat{\bs{\Sigma}}_{\text{LW}} = \left(1 - \frac{b_n^2}{d_n^2}\right) \bs{S} + \frac{b_n^2}{d_n^2}\hat{\mu} \bs{I},
\end{equation} 
where $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$, and $\Vert \cdot \Vert_F$ denotes the Frobenius norm.

@@

Our proposed estimator \eqref{proposed} is not guaranteed to be positive-definite. To correct this, we reshape our vector estimator back into a matrix and then identify the closest positive-definite matrix. \citet{higham1988computing} and \citet{huang2017calibration} showed that the projection of a $p \times p$ symmetric matrix $\bs{B}$ onto the space of positive semi-definite matrices is
\[
P_0(\bs{B})
=
\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert
=
\bs{Q}\text{diag}\{\max(\lambda_1,0),\max(\lambda_2,0),\ldots,\max(\lambda_p,0)\}\bs{Q}^\top,
\]
where $\Vert \cdot \Vert$ denotes the Frobenius norm, $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues.

To guarantee positive-definiteness, we follow \citet{huang2017calibration} and replace non-positive eigenvalues with a chosen positive value $c$ smaller than the least positive eigenvalue $\lambda_{\min}^+$, so that the corrected estimate is
\begin{equation}
  \label{near posdef}
  P_0(\bs{B})
  =
  \bs{Q}\text{diag}\{\max(\lambda_1,c),\max(\lambda_2,c),\ldots,\max(\lambda_p,c)\}\bs{Q}^\top.
\end{equation}
\citet{huang2017calibration} suggest $c_{\alpha}=10^{-\alpha}\lambda_{\min}^+$, where the parameter $\alpha$ is chosen to minimize $\Vert B - P_{c_{\alpha}}(B) \Vert + \alpha$ over a uniform partition of $\{\alpha_1,\ldots,\alpha_K\}$ of $[0,\alpha_K]$. In our implementations we used $K=20$ and $\alpha_K=10$.

@@

Finally, we have so far assumed that our data are multivariate normal, though simulations in the Supporting Information suggest that our procedures can still provide relatively accurate estimates when the normality assumption is violated. To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f(\cdot \mid \bs{\eta})$ in the nonparametric maximum compositive likelihood problem \eqref{Gnd hat} and in our proposed estimator \eqref{proposed}. If $f$ is unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.

>>>>

\begin{prop}
  \label{prop:Rhat}
  Define $\hat{\Delta}_{jk}^2 = \sum_{i=1}^{n}(X_{ij}X_{ik}-s_{jk})^2 / n^2$ and
  \begin{equation}
    \label{eq:Rhat}
    \hat{R}_L(\beta_S,\beta_I) = \frac{1}{p^2} \sum_{j,k=1}^{p}[(2\beta_S-1) \hat{\Delta}_{jk}^2 + \{(1- \beta_S) s_{jk} - \beta_I u_{jk}\}^2 ].
  \end{equation}
  If $p^{-2} \bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ is bounded as $n \rightarrow \infty$, then
  \[
    \lim_{n \rightarrow 0} \{ \bb{E} \hat{R}_L(\beta_S, \beta_I) - R(\bs{\Sigma}, \bs{\delta}) \} = 0
  \]
  for the class of decision rules $\bs{\delta}$ defined in \eqref{linear class}. $\Vert \cdot \Vert_F$ denotes the Frobenius norm $\Vert \bs{A} \Vert_F^2=\sum_{j,k=1}^p A_{jk}^2$.
\end{prop}
The condition on $\bb{E} \Vert \bs{S} - \bs{\Sigma} \Vert_F^2$ requires that the variances of the entries of $\bs{S}$ do not grow too quickly as $n$ grows. It is also required by \citet{ledoit2004well} and is implied by their Lemma 3.1.

Proposition \ref{prop:Rhat} shows that $\hat{R}_L(\beta_S, \beta_I)$ \eqref{eq:Rhat} is an asymptotically unbiased estimate of the true risk \eqref{frobenius risk} for linear decision rules of the form \eqref{linear class}. It is then reasonable to estimate the best estimator in this class by minimizing $\hat{R}_L(\beta_S, \beta_I)$ over $\beta_S$ and $\beta_I$. A slight modification of this procedure turns out to recover the \citet{ledoit2004well} estimator
\begin{equation}
\label{eq:lw}
\hat{\bs{\Sigma}}_{\text{LW}} = \left(1 - \frac{b_n^2}{d_n^2}\right) \bs{S} + \frac{b_n^2}{d_n^2}\hat{\mu} \bs{I},
\end{equation} 
where $\hat{\mu} = \text{tr}(\bs{S})/p$, $d_n^2=\Vert\bs{S}-\hat{\mu} \bs{I} \Vert_F^2$, $b_n^2 = \min(d_n^2, \sum_{i=1}^n \Vert \bs{X}_i \bs{X}_i^\top - \bs{S} \Vert_F^2 / n^2)$.

@@

Our proposed estimator \eqref{proposed} is not guaranteed to be positive-definite. To correct this, we reshape our vector estimator back into a matrix and then identify the closest positive-definite matrix. \citet{higham1988computing} and \citet{huang2017calibration} showed that the projection of a $p \times p$ symmetric matrix $\bs{B}$ onto the space of positive semi-definite matrices is
\[
P_0(\bs{B})
=
\argmin_{\bs{A}\geq 0} \Vert\bs{A}-\bs{B}\Vert_F
=
\bs{Q}\text{diag}\{\max(\lambda_1,0),\max(\lambda_2,0),\ldots,\max(\lambda_p,0)\}\bs{Q}^\top,
\]
where $\bs{Q}$ is the matrix of eigenvectors of $\bs{B}$, and $\lambda_1,\ldots,\lambda_p$ are its eigenvalues.

To guarantee positive-definiteness, we follow \citet{huang2017calibration} and replace non-positive eigenvalues with a chosen positive value $c$ smaller than the least positive eigenvalue $\lambda_{\min}^+$, so that the corrected estimate is
\begin{equation}
  \label{near posdef}
  P_c(\bs{B})
  =
  \bs{Q}\text{diag}\{\max(\lambda_1,c),\max(\lambda_2,c),\ldots,\max(\lambda_p,c)\}\bs{Q}^\top.
\end{equation}
\citet{huang2017calibration} suggest $c_{\alpha}=10^{-\alpha}\lambda_{\min}^+$, where the parameter $\alpha$ is chosen to minimize $\Vert B - P_{c_{\alpha}}(B) \Vert + \alpha$ over a uniform partition of $\{\alpha_1,\ldots,\alpha_K\}$ of $[0,\alpha_K]$. In our implementations we used $K=20$ and $\alpha_K=10$.

@@

Finally, we have so far assumed that our data are multivariate normal, though simulations in the Supporting Information suggest that our procedures can still provide relatively accurate estimates when the normality assumption is violated. To extend our procedure to non-normal data belonging to a parametric family, we can simply modify the density function $f_1$ and $f_2$ in the nonparametric maximum compositive likelihood problem \eqref{Gnd hat} and in our proposed estimator \eqref{proposed}. If $f_1$ and $f_2$ are unknown or difficult to specify, alternative procedures may be necessary to approximate the optimal separable rule.
